---
title: "Nitrogen and Precipitation Effect on Biomass"
author: "Rosalee Elting"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: true
    toc_float: true
    number_sections: true
    theme: "cerulean"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Overview
Same exercise as last week, but Beta0 isn't set, it has random error and is a linear relationship based on precipitation for that site level mean. 

## Sites
Originally I ran this for 8 sites, but then increased to 16 sites, with two duplicates of each precipitation level. 

```{r}
library(tidyverse) 
library(rstan)
library(tidybayes)

sigma_site <- 25
sigma_plot <- 50 
n_site <- 16
n_plot <- 6
beta1 <- 200

#setting an x value, we're not modeling the x value, we're setting the sites at known precipitations for each of 8 sites 
precip <- rep(c(300,350, 400,450, 500, 550, 600, 700), each =2)

beta0_mean <- 50 + 0.6*precip


beta0_j <- beta0_mean + rnorm(n_site, 0, sigma_site)
plot(precip, beta0_j)

```

# Generating Fake Data 
```{r}
biomass <- data.frame(
  site = rep(1:16, each = n_plot),
  plot = rep(1:6),
  nit = rep(c(rep(0, n_plot/2), rep(1, n_plot/2)), n_site),
  b_ij = NA
)
#Bij = Beta0,j + Beta1* Nit


#making site means using overall mean. 
#making data points using plot means
plot_mean <- rep(beta0_j, each = n_plot)


biomass$b_ij <- plot_mean + beta1*biomass$nit + rnorm(n_site*n_plot, mean= 0, sd= sigma_plot)

#testing to see if it matches 
mean(biomass$b_ij[biomass$nit==1])
mean(biomass$b_ij[biomass$nit==0])

#plotting based on site
ggplot(biomass, aes(x=as.factor(nit), y = b_ij, col= as.factor(nit)))+ 
  geom_point()+
  facet_wrap(~as.factor(site))+
  theme_classic()
```

# Writing Stan Model Now

```{r}


biomass_data <- list(n= nrow(biomass), n_site = as.integer(n_site), y= as.numeric(biomass$b_ij), site_id= biomass$site, nit = biomass$nit, precip = precip)

biomass_fit <- stan(file = "biomassnitprecip.stan", data = biomass_data, iter = 1000, chains= 4)
print(biomass_fit)
```

## Commented Stan Code 
```{r}

# data {
#   int<lower=0> n;
#   int<lower=0> n_site;
#   vector<lower=0>[n] y;
#   int<lower=0> site_id [n] ;
#   int<lower=0> nit [n];
#   vector<lower=0> [n_site] precip;
# }
# 
# parameters {
#   real<lower=0> sigma_plot;
#   real<lower=0> sigma_site;
#   real<lower=0> beta1; // effect of adding Nitrogen
#   real site_int;
#   real<lower=0> site_slope;
#   vector<lower=0> [n_site] mu_site; 
# }
# 
# model {
#   for (i in 1:n){
#     y[i] ~ normal(mu_site[site_id[i]] + beta1*nit[i], sigma_plot);
#   }
#   mu_site ~ normal(site_int + site_slope*precip, sigma_site);
#   sigma_plot ~ normal(0, 100);
#   sigma_site ~ normal(0, 100);
#   beta1 ~ normal(100,300);
#   site_int ~ normal(0,50);
#   site_slope ~ normal(0.6, 0.6);
# }

```


# Plotting Result of Model 
Boxplots and Violin plots represent the estimated density of intercepts from Stan Model. Points represent the probability distribution (95% Confidence Interval) of the fake data that I generated. At each precipitation there are two sites, hence the multiple points at each location. The data represented here are the **intercepts** of a model with a slope of 0.6. This represents the difference in biomass intercept of the linear regression, based on both nitrogen supplementation and precipitation at each site. 

```{r}
#summary of Stan model data
biomass_fit_summ<- summary(biomass_fit)

#mean calc is the summary of the generated data 
biomass_data <- data.frame(biomass_data)

mean_calc<- biomass_data %>%
  group_by(site_id) %>%
  summarise(meanB=mean(y), se=sd(y)/sqrt(6)) %>%
  mutate(precip = precip)

#steps is the result of the stan model's data
precip <- data.frame(precip,
                     site_id = c(1:16))

steps<- biomass_fit %>% 
  spread_draws(mu_site[site_id]) %>%
  inner_join(precip)


ggplot(mean_calc, aes(x=as.factor(precip), y = meanB))+
  geom_violin(data= steps, mapping = aes(x= as.factor(precip), y= mu_site))+
  geom_boxplot(data= steps, mapping = aes(x= as.factor(precip), y= mu_site), width=0.1, color="grey", alpha=0.2)+
     geom_point(data=mean_calc, mapping = aes(x=as.factor(precip), y=meanB-2.3*se), color="red", position = position_jitter(seed = 1, width = 0.2))+
  geom_point(data=mean_calc, mapping = aes(x=as.factor(precip), y=meanB), color="black", position = position_jitter(seed = 1, width = 0.2))+
   geom_point(data=mean_calc, mapping = aes(x=as.factor(precip), y=meanB+2.3*se), color="blue", position = position_jitter(seed = 1, width = 0.2))+
  xlab("Precipitation")+
 ylab("Biomass Regression Intercept")+
  theme_classic()

```




