---
title: "EM_S23_Week3"
author: "Rosalee Elting"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Zooplankton Posterior Example
1. generate data of zooplankton; 6 samples; poisson distribution. 
rpois(6, 20)
20 is true value of lambda

2. prior experiments (x3)
experiment 1: mu = 20; var = 50 
experiment 2: mu = 10; var =3
experiment 3: mu = 30; var = 10

3. Reparameterize
my prior distribution is gamma, take inputs of alpha and beta. We only have mu and var though. 
We need to re-parameterize gamma, using moment magic. 
mu = alpha/beta 
var = (alpha/beta)^2
Doing the algebra of this: 
alpha = mu^2/var
beta= mu / var

3. calculate alpha and beta for each prior

4. Plot posterior in red; prior in black for each  
gamma distribution- give it alpha and beta posteriors 
sample size is 6 


```{r}
#generate data set 
zoo <- rpois(100, 20)
lambda <- seq(from = 0, to = 40, by = 1)


#PRIOR 1
#make a summary of my prior variables, first is alpha prior, second is beta prior
prior1 <- c(8,0.4)
#create new data frame for posterior alpha and beta 
posterior1 <- NA
#compute the alpha posteriors which = sum(yi) + alpha prior
posterior1[1] <- sum(zoo) + prior1[1]
#compute the beta posteriors which = n + beta prior
posterior1[2] <- length(zoo) + prior1[2]

#now generate a gamma distribution with the lambda values I made a sequence of, the shape is alpha and rate is beta (both prior)
dpriorgamma1 <- dgamma(x=lambda, shape = prior1[1], rate = prior1[2])
#now generate a gamma distribution with the same lambda values, but where the shape and rate are the alpha and beta posterior. 
dpostgamma1 <- dgamma(x =lambda, shape = posterior1[1], rate = posterior1[2])

#PRIOR 2 
prior2 <- c(33.3, 3.33)

posterior2 <- NA
posterior2[1] <- sum(zoo) + prior2[1]
posterior2[2] <- length(zoo) + prior2[2]

dpriorgamma2 <- dgamma(lambda, prior2[1], prior2[2])
dpostgamma2 <- dgamma(lambda, posterior2[1], posterior2[2])

#PRIOR 3 
prior3 <- c(90,3)

posterior3 <- NA
posterior3[1] <- sum(zoo) + prior3[1]
posterior3[2] <- length(zoo) + prior3[2]

dpriorgamma3 <- dgamma(lambda, prior3[1], prior3[2])
dpostgamma3 <- dgamma(lambda, posterior3[1], posterior3[2])

plot(dpriorgamma1 ~lambda, type = "l", ylim = c(0,.8))
lines(dpostgamma1~lambda, col= "red")

plot(dpriorgamma2 ~lambda, type = "l", ylim = c(0,.8))
points(dpostgamma2 ~lambda, col= "red", type = 'l')

plot(dpriorgamma3~lambda, type = "l", ylim = c(0,.8))
points(dpostgamma3~lambda, col= "red", type = 'l')

#the more data you collect, the more it will shift to the likelihood than the prior. 
#the effect of the prior goes away if you use likelihood heavily. 

```

# Week 3 - Day 2
### Function for Posterior Probability 
```{r}
poispost <- function(P, alphaprior, betaprior, y) {
	yhat <- P[1] 
	#sum is positive here, we're maximizing the log likelihood
	post <- sum(dpois(y, yhat, log=TRUE))+ dgamma(yhat, alphaprior, betaprior, log=TRUE)
	return(post)
}
```

### The priors we are using, which can easily change and the alphaprior and betaprior are standardized formulas  
```{r}
mu <- 20 
v <- 50 
#need to make into our parameters with moment matching
alphaprior <- mu^2/v
betaprior <- mu/v
```

### Testing running the function with some sample data to make sure the fucntion works. 
```{r}
y <- c(20,19)
P = 18

poispost(P, alphaprior, betaprior, y=y)
# this result is proportional to the probability, but it's not EQUAL to it, we need a normalization constant to change from proportional to equal to probability. 
#expected value here is -7.77
```
### The data we will use is defined here
```{r}
#the data set we're working with from now on 
y <- rpois(6,20)
```

#Running our own MCMC chain
```{r}
#make a matrix, called chain, to store our parameters in. we're taking 10000 steps, and we're using 1 column for the lambda, but can make larger later. 
chain <- matrix(nrow= 10000, ncol= 1)
#don't worrry about the size of the scale, we'll adapt later.
scale <- 5
#start the parameter estimate somewhere, here we start it at 1. 
chain[1,] <- c(1)

for(i in 2:nrow(chain)) {
	positive <- F
	while(positive==F){
	pstar <- rnorm(1, chain[i-1,], scale)
	positive <- ifelse(pstar >1, T, F)}
	poststar <- poispost(pstar, y=y, alphaprior=alphaprior, betaprior=betaprior)
	postchain <- poispost(chain[i-1], y=y, alphaprior=alphaprior, betaprior=betaprior)
	R <- exp(poststar - postchain)
	rprob <- runif(1)
	chain[i,] <- ifelse(R > rprob, pstar, chain[i-1]) 
}

#while loop only for Poisson 
```
### Checking acceptance and plotting
```{r}
#and we calculate acceptance. If we accept none, it's a bad estimate we started with, if we accept them all, we're taking too small of steps. We aim for 0.2. 

acceptance <- 1- mean(duplicated(chain[,1]))
print(acceptance)

#then we'll plot it 
plot(chain[,1], type = "l")

#plotting the posterior probability density; the integral of this is 1. 
hist(chain[,1])
#or
plot(density(chain[,1]))
```
### Reporting some stats on this model 
```{r}
median(chain[,1])
mean(chain[,1])
var(chain[,1])
quantile(chain[,1], c(.025, 0.5, 0.975))

#the difference between the .025 and 0.975 is our 95% uncertainty interval or 95% credible interval. 
# we're 95% confident that the true lambda of our data is between those two values. 
```
### Comparing with just the priors and posteriors from our actual data's mu and var
```{r}
#mu = alpha/beta
#var = alpha/beta^2
qgamma(0.025, 122, 6.4)
qgamma(0.975, 122, 6.4)
# the approximation will never match perfectly. 
```

