---
title: "EM_S23_Week4"
author: "Rosalee Elting"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Making fake data 
```{r, echo=FALSE}
slope <- 1
intercept <- 5
error <- rnorm(n=20, mean =0, sd =3)
x <- seq(1:20)
y <- intercept + slope* x + error
plot(x,y)
```

# Setting up a list 
Stan takes a list, a compilation of objects. We don't want a df, because it has different size input. So we need to be able to put variable amounts of data into Stan. 

```{r}
#first thing is what needs to list the stan file (in this case it's N). then inputting data from randomly generated data above.  

reg_data <- list(N= length(x), x= x, y = y )
```

# Running the stan model on my data 

```{r}
library(rstan)
reg_fit <- stan(file = "linreg.stan", data = reg_data, iter = 1000, chains= 4)
print(reg_fit)
```
Last output is log posterior, it's the log of the posterior probability at values of different parameters. 
sd is standard deviation of the posterior probability distribution 
```{r}
plot(reg_fit)
traceplot(reg_fit)
```
Our chains produce less auto-correlation within a chain. Looks way better than the one we made in the MCMC homework. It's already more powerful with 1000 steps than 100k in our own code. Shaves a lot of time with less samples, they are musch better behaved than the other one (due to Hamiltonian physics, stan is making much better proposals).

Rhat combines chain stationarity with multiple chains. High R hat corresponds to two nice chains that have very different estimates. Takes the stationarity of multiple chains. Way to beat this is run each chain longer (if they aren't the same shape) or more chains (if they are stable but the estimates are different). ALWAYS LOOK AT YOUR CHAINS TO CHECK THIS. Problems with computation means problems with your model. 

Rule of thumb is r-hat should be < 1.1. Happy place are stationary chains that are mixing right on top of eachother. 

# Compare with canned function 
Here comparing with a canned lm function for our data. 
```{r}
summary(lm(y~x))
```
```{r}
library(shinystan)
launch_shinystan(reg_fit)
```

Tidy base uses language of dplyr to do things. Here we make a data frame of each chain with our step and a and b values. 

What we have is the steps in the chain for each parameter. This matrix of steps is the join distribution, the marginal is any one of these.
```{r}
library(tidyverse)
library(tidybayes)

regsteps <- reg_fit %>% 
  spread_draws(a , b)
```

We can calculate the joint distribution for the line for each one of these. We're going to plot the relationship of this joint distribution for each step in the chain. This is the model error, not the prediction error. 

```{r}
plot(x,y, xlim=c(0,22), ylim= c(0,30))

for(i in 1: 1000) { 
  #this is my model uncertainty. gets smaller with more samples. 
  lines(x, regsteps$b[i] * x + regsteps$a[i], col = alpha("lightblue", 0.1))
  #to make a data set to include the prediction error . shows prediction uncertainty, this doesn't change with higher n. 
  points(x, regsteps$b[i] * x + regsteps$a[i]+ rnorm(1, 0, 3.6), pch = 16, col= alpha("darkgray", 0.05))
}
points(x, y, pch=16, col = "darkblue")
lines(x, x*mean(regsteps$b) + mean(regsteps$a), col = "gray8")
```


# Week 4 - Day 2 
Biomass example making sample data. Have to start at the top with the mean of everything and then work our way down. 
Here is some data we're making up. 
```{r}
#given data
set.seed(40)
sigma_site <- 50 
sigma_plot <- 50 
n_site <- 8 
n_plot <- 6

b_mean <- 400
```


Generating Means of the data from 6 plots in 8 sites 
```{r}
#making site means using overall mean. 
site_mean <- rep(b_mean, n_site) + rnorm(n_site, 0, sigma_site)
#making data points using plot means
plot_mean <- rep(site_mean, each = n_plot) + rnorm(n_plot*n_site, 0, sigma_plot)

fake_data <- data.frame(
              site = rep(1:8, each = n_plot), 
              b_ij = plot_mean)
#testing to see if it matches 
mean(fake_data$b_ij[fake_data$site==6])
#plotting based on site
plot(x= fake_data$site, y = fake_data$b_ij)

```
# Running the Stan Model 
```{r}
library(rstan)

biomass_data <- list(N= nrow(fake_data), N_site = as.integer(n_site), y= as.numeric(fake_data$b_ij), site_id= fake_data$site)

biomass_fit <- stan(file = "biomassex.stan", data = biomass_data, iter = 1000, chains= 4)
print(biomass_fit)
```
```{r}


```


