---
title: "EM_S23_Week6"
author: "Rosalee Elting"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Same exercise as last week, but Beta0 isn't set, it has random error and is a linear relationship based on precipitation for that site level mean. 

```{r}
sigma_site <- 25
sigma_plot <- 50 
n_site <- 8
n_plot <- 6
beta1 <- 200

#setting an x value, we're not modeling the x value, we're setting the sites at known precipitations for each of 8 sites 
precip <- c(300,350, 400,450, 500, 550, 600, 700)

beta0_mean <- 50 + 0.6*precip
 plot(precip, beta0_mean)


beta0_j <- beta0_mean + rnorm(n_site, 0, sigma_site)
plot(precip, beta0_j)

```
```{r}
biomass <- data.frame(
  site = rep(1:8, each = n_plot),
  plot = rep(1:6),
  nit = rep(c(rep(0, n_plot/2), rep(1, n_plot/2)), n_site),
  b_ij = NA
)
#Bij = Beta0,j + Beta1* Nit


#making site means using overall mean. 
#making data points using plot means
plot_mean <- rep(beta0_j, each = n_plot)


biomass$b_ij <- plot_mean + beta1*biomass$nit + rnorm(n_site*n_plot, mean= 0, sd= sigma_plot)

#testing to see if it matches 
mean(biomass$b_ij[biomass$nit==1])
mean(biomass$b_ij[biomass$nit==0])

#plotting based on site
ggplot(biomass, aes(x=as.factor(nit), y = b_ij, col= as.factor(nit)))+ 
  geom_point()+
  facet_wrap(~as.factor(site))+
  theme_classic()
```

# Writing Stan Model Now

```{r}
library(rstan)

biomass_data <- list(n= nrow(biomass), n_site = as.integer(n_site), y= as.numeric(biomass$b_ij), site_id= biomass$site, nit = biomass$nit, precip = precip)

biomass_fit <- stan(file = "biomassnitprecip.stan", data = biomass_data, iter = 1000, chains= 4)
print(biomass_fit)
```

Make this for 16 sites instead. Will need to rep each precip twice. rep(precip, each =2)
Plot 1: plot level model (facets and lines), facet (.~site), and have a line for each. Mean line of the intercept (site dependent) and the slope (0.61 estimated from)- Won't do because of time. 

plot 2: x = precip, y = steps and plot posteriors using spread_draws
  mu_site is the site level intercept (mu_site[6] isn Stan. Do violin or boxplot. 


