---
title: "EM_S23_Week2"
author: "Rosalee Elting"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Week 2 - Day 1
### Poisson distribution of discrete variables 
Inputs: 
Lambda 
some z 

Goal : plot probability mass function for lambda = 4 
calculate mean 
calculate variance 
sum [z]

```{r}
l <- 4
z <- seq(from = 0,to = 20,by =1)
  

pois <- function(z,l){ 
  zgivl <- ((l^z)/factorial(z))*exp(-1*l)
}

zgivl <- pois(z,l)

mu <- sum(z*zgivl)
sig <- (z-mu)^2*zgivl
add <- sum((z-mu)^2*zgivl)

plot(z, zgivl)

#the mean is not at 4 because there is an infinite amount of option for counts we could take. 
```
### Binomial distribution 
```{r}
z <- seq(from =0, to =10, by =1)
phi = 0.4
n <- 10
nz <- factorial(n)/(factorial(z)*factorial(n-z))

bino <- function(n, z, phi){
  probz <- nz*(phi^z)*((1-phi)^(n-z))
}

zgivnphi <- bino(n, z, phi)

#mean 
mu <- sum(z*zgivnphi)
sig <- (z-mu)^2*zgivnphi
#variance
add <- sum((z-mu)^2*zgivnphi)

plot(z, zgivnphi)

```
### Looking at real data from a paper 
This is fake data! 
```{r}
fish <- read.csv("fish_choice.csv")
phi = 0.5

fish <- as.numeric(fish$n)
mu <- mean(fish)
sig <- phi*24*(1-phi)

add <- sum(phi*24*(1-phi))
 
```

```{r}
#we're making a data set 
# of randoms 
#input for random binomial (number of repetitions, number of trials, phi probability of success)

rand <- rbinom(20, 24, 0.5)
hist(rand)
hist(fish)
# if wanted to make a loop of many of these random data sets and ordering their variances to see if we can get the 0.6 variance these authors report. 

varfish <- NA
for (i in 1:10000) { 
  varfish[i] <- var(rbinom(20,24,0.5))
}
hist(varfish)
sort(varfish, decreasing = FALSE)

```
# Week 2 - Day 2 
### Plotting a Normal Distribution 
Moments are the parameters 
the moments in a normal are the parameter 

```{r}
#Given the following variables. Chose z as it centers around the moment, which in a normal distribution is mu
mu <- 10 
sig <- 5 
z <- seq( from = -10, to = 30, by =1)

norm <- dnorm(z, mu, sig)
plot(z, norm, col = "darkblue", ylim = c(0, 0.1), type = "l")
```
### Poisson with probability 
```{r}
#making a poisson if we know a data point and the lambda
pois <- dpois( x= 14, lambda = 20)
```
### Poisson with likelihood
```{r}
#plotting a likelihood with only one known point for several possible lambdas
#assumes your data are Poisson distributed
#searching across the parameters to see what is most likley 

lambda <- seq(from = 1, to = 20, by =1)
like <- dpois( x= 6, lambda = lambda)

plot(lambda, like)
```

```{r}
#the joint probability of drawing sample 1, sample 2 (6 and then 8)
lambda <- seq(from = 1, to = 20, by =1)
like6 <- dpois( x= 6, lambda = lambda)
like8 <- dpois (x=8, lambda = lambda)

prod <- like6*like8

plot(lambda, like6, col= "red")
plot(lambda, prod, col = "blue")

```
### Log likelihoods 
generate 10 random data points from poisson
lambda = 20 
estimate hat-lambda using likelihood profile 
loop through lambda 

```{r}
x<- rpois(n= 100, lambda = 20)
lambda <- seq(from = 15, to = 25, by =.25)
prob <- NA

for (i in 1:length(lambda)) {
  prob[i] <- sum(dpois(x = x, lambda= lambda[i], log=T))
}

#limit plot to be able to see the max and guesstimate the confidence intervals 
plot(lambda, prob, xlim = c(18,22), ylim = c(-300, -280))
```

### Plotting Likelihoods with a known
Generate random normal data 
rnorm(10, mean =10, sd =3)
estimate likelihood profile given that theta = 3 (which is the standard deviation; sigma; stdev)


```{r}
x <- rnorm(10, mean=10, sd =3)

mu <- seq(1, 20, 1)
sigma <- 3
#sigma is variance
prob <- NA

for (i in 1:length(mu)) {
  prob[i] <- sum(dnorm(x = x, mean = mu[i], sd= 3, log=T))
}

plot(mu, prob, ylim = c(-30, -20), xlim = c(7,15))
```
### Non Linear Minimization
write a function to return the negative log likelihood of a poisson distribution 
```{r}
#making our own pois like function, later in the nln as the function to minimize

# first try to get used to the function
#works
poislike <- function(data, lambda){
 y<- -sum(dpois(x = x, lambda= lambda, log=T))
}

```
```{r}
#P is vector of parameters
#yhat is lambda values 
#this returns the negative log likelihood

poislik <- function (P, data) { 
  yhat <- P[1]
  lik <- -sum(dpois(x= data, lambda = yhat, log=T))
  return(lik)
}

#setting data 
data <- rpois(n= 100, lambda = 20)
#non linear minimization
nlm(poislik, p = c(1), data =data)
#we're interested in the estimate and minimum (which is of the negative minimum log likelihood)
#can check by comparing $estimate from nlm and the mean(data), should be very close. 
mean(data)
```
### Linear Likelihood 
where:
m =1 
b = 5 
sigma/sd = 3
n=20
x= seq(from = 6, to =25, by =1)

```{r}
# generate a random data set with these variables 
set.seed(30)
x <- seq(from = 6, to =25, by =1)
e <- rnorm(n=20, mean=0, sd =3)
y <- 1*x+5+e

reglike <- function(P, y, x) { 
    yhat <- P[1]*x +P[2]
    lik <- -sum(dnorm(x=y, mean= yhat, sd = exp(P[3]), log=T))
    return(lik)
} 

nlm(reglike, p = c(1,2,3), y, x)
#returns the log of P[3]
#so need to exponentiate the P[3] again from the above code. It is the third value in $estimate
exp(1.0407128)

plot(y)
summary(lm(y~x))
```


