---
title: "EM_S23_Week1"
author: "Rosalee Elting"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Week 1 - Day 1
### Writing a Michaelis-Menten Function 
Fixed Values: 
Umax = 1 
Ks = 100 
Formula U = Umax S / Ks + S
```{r mm-function}
mm_function <- function(Um, Ks, S){
 U <- Um*S/(Ks+S)
 plot(S,U)
 }
S <- seq(from = 0, to = 200, by = 1)
mm_function(Um =1, Ks = 100, S = S)
```
### Using a Threshold Switching Equation
x = sequence from 1 to 30 
Tao = 17 
g1 y = b
g2 y = ax+b 
b = 10 
a = 5

Here is the structure of the argument I should use. 
ifelse(condition- a Boolean statement , value 1, value 2)
The condition: X is <17 or >=17 

```{r threshold-switching}
x <- seq(1:30)
a <- 5
b <- 10 
y <- ifelse(x >= 17, a*x+b, b)

plot(x, y)
```
### Week 1 - Day 1 Homework 
#### For Loops
Gotelli - Basic Anatomy and Use of For Loops

for(var in seq) { 
Body for the loop, often the call to a function 
}

var is a counter variable that holds the current values of the counter in the loop
Seq is an integer vector (defines start and end values of the loop)

i,j,k for var (counter)

Using seq_along(df) is a safe way to do rather than length(df) so that it won't retun a result if vector is blank. 

Loop tips: 
Do NOT change object dimenstion inside a loop 
Avoid these functions: cbind, rbind, c, list


```{r}
my_dogs <- (c("chow", "akita", "malamute", "husky", "samoyed"))

for(i in seq_along(my_dogs)) {
  cat("i =", i, "my_dogs[i] =", my_dogs[i], "\n") 
}
```
```{r}
my_bad_dogs <- NULL
for (i in seq_along(my_bad_dogs)) {
  cat("i =", i, "my_bad_dogs[i] =", my_bad_dogs[i], "\n")
}
```
Can define vector length from a constant 
```{r}
zz <- 5
for( i in seq_len(zz)) { 
  cat("i =", i, "my_dogs[i] =", my_dogs[i], "\n")
  }
```
#### If Else Function 
Format: 
ifelse(test, yes, no)
test is an object that can be coerced into a logical true/false for each element in the vector 
Yes returns values for true elements in the test 
no returns values for fasle elements in the test 

Examples: suppose we have an insect pop in which each female lays on average 10.2 eggs, following a Poisson distribution. lambda = 10.2 here. However, there is a 35% chance of parasatism, in which case no eggs are laid. Here is a random sample of eggs laid for a group of 1000 individuals. 

```{r}
#makes 1000 random uniform values 
tester <- runif(1000)
eggs <- ifelse(tester > 0.35, rpois(n=1000, lambda=10.2), 0)
hist(eggs)
```
This produces a mixture model where there is a peak at ~10.2. Also a spike at zero, which represents the 0s created from Poisson itself, but also the 35% of individuals that didn't reproduce due to parasitism. 

Example 2: 
Suppose we have a vector of probability values (such as from a simulation). We want to highlight significant values in the vector for plotting 

```{r}
p_vals <- runif(1000)
z <- ifelse(p_vals <=0.025, "lower_tail", "non_significant")
z[p_vals >=0.975] <- "upper_tail"

table(z)
```

# Week 1 - Day 2
### Dyanmic Models - Exponential growth model 
Formula: 
Nt = N0 e^kt
Known values: 
K = 0.05 
N0 =10 
```{r}
#as a function
dyn_function <- function(K,N0){ 
 Nt <- N0*exp(K*t)
 plot(t,Nt)
}
t <- seq(from = 0, to = 200, by = 0.5)
dyn_function(K=0.05, N0= 10)

#not as a function, as a single vector. 
N0 <- 10 
K <- .05
t <- seq(from =0, to = 200, by = 0.5)
Nt <- N0*exp(K*t)
plot(t, Nt)
```
### Ricker Model 
Formula: 
Nt+1 = Nt*e(rmax(1-N/K))
Known values: 
k = 100 
r = 1.5
t = 0 to 100
We're forced to write a loop, because the current value is the input for the following value. Need a starting value, so we'll use: 
N1= 60 

```{r}
#t <- seq(0:100)
k <- 100
r <- 3.0
N <- NA
N[1] <- 60
t <- seq(1:100)

#the below works, but I also want to standardize the way I assigned the length so it's not fixed in the loop but rather pre-assigned. 

# for(i in 2:100) {
#   N[i] <- N[i-1]*exp(r*(1-N[i-1]/k))
# }

for(i in 2:length(t-1)) {
  N[i] <- N[i-1]*exp(r*(1-N[i-1]/k))
}
#plotting N over time 
plot(t, N, type = "l", ylim = c(0,300))
     
#plotting Nt vs Nt-1 as a plot. This is following a pattern and deterministic, even though prior graph looks like pure chaos. 
# Phase-space diagram 
plot(N[1:99], N[2:100])
```
### Numerical Approximation 
Formulas we're using: 
Exponential growth
Nt = N0e^kt
Numerical approximation: 
Nt = Nt-1 KNt-1 deltaT
Known values: 
K = 0.01
N0 =1 
up to t =20 
Goal: plot both on same axis, use different colors. 

```{r}
K <- 0.1 

NNt <- NA
NNt[1] <- 1
N0 <- 1
t <- seq(from =1, to = 20, by = 1)
#exponential 
ENt <- N0*exp(K*t)
#numerical integration 
for(i in 2:length(t-1)) {
  NNt[i] <- NNt[i-1]+K*NNt[i-1]*1 
}

#N[i] <- N[i-1]*exp(r*(1-N[i-1]/k))
plot(t, ENt, type= "line", col = "violet",ylab = "Nt")
lines(t, NNt, col= "cornflowerblue")

```
### Numerical Integration of Lotka- Volterra Model 
Formulas: 
Nt = Nt-1 + (rNt-1 - aNt-1Pt-1) delta t
Pt = Pt-1+ (abNt-1 Pt-1  - sPt-1) delta t
Known values : 
delta t = 0.1 
length of run = 2000
n[1] = 20 
p[1] = 1 
r = 0.04
a = 0.005
s= 0.1

```{r}
n <- seq(from = 1, to = 200.9, by =0.1)
Nt <- NA
Nt[1] <- 20
Pt <- NA
Pt[1] <- 1
r <- 0.04
a <- 0.005 
b <- 0.2
s <- 0.1
deltat <- 0.1

for (i in 2:length(n)) { 
  Nt[i] <- Nt[i-1] + (r*Nt[i-1] - a*Nt[i-1]*Pt[i-1])*(n[i]-n[i-1])
  Pt[i] <- Pt[i-1] + (a*b*Nt[i-1]*Pt[i-1]- s*Pt[i-1])*(n[i]-n[i-1])
}

plot(n, Nt, col = "violet",ylab = "Nt")
lines(n, Pt, col= "darkblue")
#phase diagram is a circle 
#there is an equilibrium that exists in the center of this circle. As N goes up, then P catches, and raises until N decreases and P does again. 
plot(Nt, Pt)
```

 




