---
title: "Group Level Models"
author: "Rosalee Elting"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: true
    toc_float: true
    number_sections: true
    theme: "cerulean"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Group level Models    
This week we are replicating the work we did last week, but including a treatment where 3 plots at each site receive Nitrogen supplement, which boosts their mean biomass by 200 grams, on top of the overall mean biomass of 400. 

# Grassland Example
For our example, we are studying biomass of Montana grasslands. This is done at 8 sites where conditions are expected to differ and have effects on biomass. Within each site, there are 6 plots, therefore a multilevel model is appropriate to represent our data. \

Our model will use the following relationship: 
$$B_{i,j} = B_j + \epsilon_{i,j}$$ 
where: 
$$\epsilon_j \sim normal(0, \sigma_{site})$$
In words this says: the grass biomass for the whole study (plots are $i$ and sites are $j$) are a result of the average biomass from each site ($B_j$) with normally distributed error. 

The prior probability for mean site biomass is represented as: 
$$ B_{j} = B_{mean} + e_{j}$$

# Load Libraries and Input Known Values

```{r}
library(tidyverse)
library(MetBrewer)
library(rstan)
library(tidybayes)
library(dplyr)

sigma_site <- 50
sigma_plot <- 50 
n_site <- 8
n_plot <- 6
beta0_mean <- 400 
beta1 <- 200
#going to need another column in df called nit, that is repeat of 0 0 0 1 1 1 at each site. 
```

# Generate Fake Dataset 
```{r}
biomass <- data.frame(
  site = rep(1:8, each = n_plot),
  nit = rep(c(rep(0, n_plot/2), rep(1, n_plot/2)), n_site),
  b_ij = NA
)
#Bij = Beta0,j + Beta1* Nit


#making site means using overall mean. 
site_mean <- rep(beta0_mean, n_site) + rnorm(n_site, 0, sigma_site)
#making data points using plot means
plot_mean <- rep(site_mean, each = n_plot) + rnorm(n_plot*n_site, 0, sigma_plot)

biomass$b_ij <- plot_mean + beta1*biomass$nit

#testing to see if it matches 
mean(biomass$b_ij[biomass$nit==1])
mean(biomass$b_ij[biomass$nit==0])

#plotting based on site
ggplot(biomass, aes(x=as.factor(nit), y = b_ij, col= as.factor(nit)))+ 
  geom_point()+
  facet_wrap(~as.factor(site))+
  theme_classic()

```

# Running Stan Code 
```{r}
library(rstan)

biomass_data <- list(n= nrow(biomass), n_site = as.integer(n_site), y= as.numeric(biomass$b_ij), site_id= biomass$site, nit = biomass$nit, beta1= beta1)

biomass_fit <- stan(file = "biomass_wnit.stan", data = biomass_data, iter = 1000, chains= 4)
print(biomass_fit)
```

# Commented Stan Code
Below is the stan code I've written, commented out but to have it in one location. 
```{r}
# data { 
#   int<lower=0> n; 
#   int<lower=0> n_site;
#   int<lower=0> beta1; 
#   vector<lower=0>  [n] y; 
#   int<lower=0>  site_id [n]; 
#   int<lower=0> nit [n];
# }
#   
# parameters { 
#   vector<lower=0> [n_site] mu_site_nit; //site level mean, with nit addition, length of 8
#   vector<lower=0> [n_site] mu_site_cntr; //site level mean, without nit addition, length of 8
#   real<lower=0> sigma_plot; //given value
#   real<lower=0>  sigma_site; //given value
#   real<lower=0> mu; //group level mean  for all treatments
# 
# }
# 
# 
#   
#   model { 
#     for( i in 1: n){
#       if (nit[i] == 1)
#           y[i] ~ normal(mu_site_nit[site_id[i]], sigma_plot);
#       else
#           y[i] ~normal(mu_site_cntr[site_id[i]], sigma_plot);
# }
#     mu_site_nit~ normal((mu + beta1), sigma_site); //eight means for each site's 3 plots with nitrogen
#     mu_site_cntr~ normal(mu, sigma_site); //eight means for each site's 3 plots without nitrogen 
#     mu ~ gamma(1, 0.001);  //this and subsequent rows are prior distribtuions for the parameters 
#     sigma_plot ~ normal(0,100); 
#     sigma_site ~ normal(0,100);
# }

```

# Visualizations 
```{r}
biomass_fit_summ<- summary(biomass_fit)

biomass.data <- data.frame(biomass_data)

mean_calc<- biomass.data %>%
  group_by(site_id, nit) %>%
  summarise(meanB=mean(y), se=sd(y)/sqrt(6))

# steps<- biomass_fit %>% 
#   spread_draws(mu_site)

steps_nit <- biomass_fit %>%
  spread_draws(mu_site_nit[site_id]) %>% 
  mutate(nit = 1 ) %>% 
  rename(mu = mu_site_nit)

steps_cntr <- biomass_fit %>%
  spread_draws(mu_site_cntr[site_id]) %>%
  mutate(nit =0 ) %>% 
  rename(mu = mu_site_cntr)

steps_comb <- steps_cntr %>%
  full_join(steps_nit)
  

ggplot(steps_comb, aes(x=as.factor(site_id), y = mu)) + 
  geom_violin()+
  geom_boxplot(width=0.1, color="grey", alpha=0.2) +
  geom_point(data=biomass.data, mapping = aes(x=as.factor(site_id), y=y),  position = position_jitter(seed = 1, width = 0.2))+
  geom_point(data=mean_calc, mapping = aes(x=as.factor(site_id), y=meanB), color= "#3399FF")+
  xlab("Site")+
  ylab("Biomass")+
  facet_wrap(.~nit)+
  theme_classic()
```

### Visualization 2
```{r}
ggplot(steps_comb, aes(x=as.factor(site_id), y=mu)) + 
  geom_boxplot(width=0.1, color="grey", alpha=0.5, outlier.shape = NA) +
  #geom_point(data=B, mapping = aes(x=as.factor(site), y=B_ij ),  position = position_jitter(seed = 1, width = 0.2))+
  geom_point(data=mean_calc, mapping = aes(x=as.factor(site_id), y=meanB), color="#3399FF")+
   geom_point(data=mean_calc, mapping = aes(x=as.factor(site_id), y=meanB+2.3*se), color="#3399FF")+
   geom_point(data=mean_calc, mapping = aes(x=as.factor(site_id), y=meanB-2.3*se), color="#3399FF")+
  xlab("Site")+
  ylab("Biomass")+
  theme_classic()+ 
  facet_wrap(.~nit)
```

