---
title: "Predator_Prey_HW_Elting"
author: "Rosalee Elting"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: true
    toc_float: true
    number_sections: true
    theme: "cerulean"
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Predator-Prey Models: Introduction 
A more realistic model beyond the Lotka-Volterra Model, which allows for a limit on the L-V model's shortcoming with infinite predator consumption when prey level are high. Here, we incorporate more variables, indicated in the variable index below by an asterisk. 

# Index of Variables in Model 

$P$ - Predator number\
$N$ - Prey number\
$r$ - growth rate of prey\
$K$ - Carrying capacity or prey\
$u_{max}$ - maximum predation rate *\
$k_s$ - density of prey where Predator (P) eat 50% of their maximum rate *\
$b$ - conversion efficiency of prey biomass to predator biomass\
$d$ - predator death rate\


# Logistic Prey Model with a functional response 
Following are two models, representing the change in prey and predators over time. They are as follows. 
## Prey Model 
$$\frac{dN}{dt}=rN(1-\frac{N}{K}) - \frac{u_{max}N}{k_s+N}\times P$$

## Predatory Model 
$$\frac{dP}{dt}= \frac{u_{max}N}{k_s+N}\times Pb - dP$$

# Numerical Integration of the Logistic Prey Models 
Below are the modified version of the above models that were transformed to represent numerical integration of the formulas for the prey and predator formulas. 
## Prey Formula 
Need to double check below, they focusing on formatting
$$N_t = N_{t-1}+(rN_{t-1}\times (1- \frac{N}{K}) - \frac {U_{max}N_{t-1}} {(K_s+N_{t-1})} \times P_{t-1}) \Delta t  $$
Nt = Nt-1  ( r*Nt-1 ( 1- (N/K)) - Umax*Nt-1 / (Ks +Nt-1))deltat
## Predator Formula
Pt= Pt-1 + ((umax*Nt-1/(Ks+Nt-1) * b * Pt-1) - dPt-1)*deltat 
$$P_t = P_{t-1} + (\frac {U_{max}N_{t-1}}{(K_s + N_{t-1})} \times bP_{t-1} - dP_{t-1}) \Delta t$$
# Execution of the Numberical Integration of the Logistic Prey Model 
Now that I have a numerical integration of this model, I will use this model to represent prey and predator relationships with the following given values. 
```{r}
#length of run : 15000
n <- seq(from = 1, t =1500.1, by = 0.1)
Nt <- NA
Pt <- NA
#Delta time, time steps
dt <- 0.1
#starting values for each data set 
Nt[1] <- 20 
Pt[1] <- 1
r <- 0.15 

b <- 0.2
d <- 0.01

K <- 25
ks <- 20 
umax <- 0.2

#Loop
for (i in 2: length(n)) {
  Nt[i] <- Nt[i-1] + (r*Nt[i-1]*(1-(Nt[i-1]/k))-(umax*Nt[i-1]/(ks + Nt[i-1]))*Pt[i-1])*(n[i]-n[i-1])
  Pt[i] <- Pt[i-1] + (umax*Nt[i-1]/(ks +Nt[i-1])*b*Pt[i-1] -d*Pt[i-1])*(n[i]-n[i-1])
}
```


# Plots of Relationships 
plot predators and prey as function of time 
and predator vs prey 
```{r}
plot(n, Nt, col = "violet", ylab = "Nt")
lines(n, Pt, col= "darkblue")

plot(Nt,Pt)
```

