---
title: "RegressionMLE_HW_Elting: Using Maximum Likelihood to Solve Regression Parameters"
author: "Rosalee Elting"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: true
    toc_float: true
    number_sections: true
    theme: "cerulean"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Michaelis-Menten Non-Linear regression

Here we will be using the non-linear Michaelis-Menten regression to complete our model. 
$$ U = \frac {U{max}S}{K{s}+S}$$
# Our Known Parameters

Known Parameters: 

$U{max}$ =1\
$K{s}$ = 50\
$\sigma$ = 0.05\
$n$=20\

# Creating the Model 
## Generating fake data

Generate fake data , assuming normally distributed errors 
*here I am generating fake data, that is normally distribute to represent my x and y values*
First I'm inputting my known values from above into R
```{r}
umax <- 1
ks <- 50 
sig <- 0.05
```

```{r}
# I want my variable S in the Michaelis-Menten formula to go from 5 to 100. 
set.seed(40)
#making a range of values that s could be pulled from. 
range <- seq(from=5, to =100, by =1)
#taking a subset of x values for s, 
s <- sample(x= range, size =20)
#making a random variable, e, to represent error points of this normal distributed data set. 
e <- rnorm(n=20, mean=0, sd =0.05)
#now inputting the M-M formula to produce u values from these s values 
u <- (umax*s/(ks+s))+e
#Plotting my fake data that follows a M-M curve. 
plot(s,u)
```

# Creating a function to report likelihood 
I will now create a function to report the likelihood of the parameters, given the data I created in the above chunk using M-M formula and randomly distributed samples.
*I do this first by creating a function*

```{r}
mmlike <- function(P, u, s) { 
  uhat <- P[1]*s /(P[2]+s)
  like <- -sum(dnorm(x=u, mean = uhat, sd = P[3], log =T))
  return(like)
}
#P[1] equals Umax, P[2] equals Ks and P[3] equals the given sigma standard deviation
```

# Searching for the best parameters 
Here I use the nlm (non linear minimization) function to estimate the best value of my parameters that will minimize the negative log likelihood of several parameters at one time. I use the mmlike function I made in the previous chunk as the function_to_minimize, one of the arguments in the nlm function.  
```{r, warning=FALSE}
mmout <- nlm(mmlike, p= c(1,2,3), u, s)
mmout

#P[3] is still in log format, so it must be exponentiation: 
#exp(-2.8722265)
```
**This returns the following parameter estimates:** \
$U{max}$ = 0.8646424 \
$K{s}$ = 38.6109660 \
$\sigma$ = 0.05657293\

# Comparing with a pre-made regression routine
Now I'll compare my estimates with the mathematical solution that can be produced with just a summary using a regression model with the two variables in R.  
```{r}
#summary()
```

# Plot model fit on the data 
Now plotting how our model fits the data I generated early in this assignment. 
```{r}
#making a line for my Michaelis-Menten model 
mmm <- 0.865*range/(38.61+range)
plot(range, mmm, type ='l', lwd = 3, col = "darkblue")
points(s,u)

```





