---
title: "Segre_data_summary"
author: "Rosalee Elting"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dplyr)
library(rstatix)
library(ggpubr)
```
# Load data 
```{r}
annas <- read_csv("annasairdata.csv")
print(annas)
```

# Cleaning Annas Data 
Only using data from a competition. 
Selecting data I am interested in from original data.

```{r}
annas_clean <- annas %>%
  filter(solocomp == "comp") %>%
  select(c("mass", "massSTD", "w.lifted", "w.liftedSTD", "wing.ar", "wing.arSTD", "wing.length", "wing.lengthSTD", "mean.hor_acel", "mean.hor_decel", "mean.arc_avghvel", "mean.total_vel", "ss.pitch_roll", "mean.pr_time", "mean.pr_degrees", "mean.arc_force", "PRTpercent"))

annas_mean <-  as.data.frame(annas_clean) %>%
  summarise(across(colnames(annas_clean), ~mean(.x, na.rm=TRUE))) %>%
  pivot_longer(cols = colnames(annas_clean)) %>%
  select(-c("name"))
annas_var <- as.data.frame(annas_clean) %>%
  summarise(across(colnames(annas_clean), ~var(.x, na.rm = TRUE))) %>%
  pivot_longer(cols = colnames(annas_clean)) %>%
  select(-c("name"))
 
annas_summ <- data.frame(matrix( NA, 
                                 ncol = 3, 
                                 nrow = 17))

names <- data.frame(rownames = colnames(annas_clean))

annas_summ[,1] <- names
annas_summ[,2] <- annas_mean
annas_summ[,3] <- annas_var
names(annas_summ) <- c("variable", "mean", "var")

```

# Plotting Distribution of Original Data 
```{r}

annas %>%
  filter(solocomp== "comp") %>%
  ggplot(aes(x= as.factor(0), y= mass))+
  geom_violin()+
  geom_point()+
  theme_classic()

annas %>%
  filter(solocomp== "comp") %>%
  ggplot(aes(x= as.factor(0), y= wing.length))+
  geom_violin()+
  geom_point()+
  theme_classic()

annas %>%
  filter(solocomp== "comp") %>%
  ggplot(aes(x= as.factor(0), y= wing.ar))+
  geom_violin()+
  geom_point()+
  theme_classic()

annas %>%
   filter(solocomp== "comp") %>%
  ggplot(aes(x= as.factor(0), y=w.lifted))+
  geom_violin()+
  geom_point()+
  theme_classic()

annas %>%
   filter(solocomp== "comp") %>%
  ggplot(aes(x= as.factor(0), y=PRTpercent))+
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75))+
  geom_point()+
  theme_classic()

annas %>%
   filter(solocomp== "comp") %>%
  ggplot(aes(x= as.factor(0), y= mean.hor_acel))+
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75))+
  geom_point()+
  theme_classic()

annas %>%
   filter(solocomp== "comp") %>%
  ggplot(aes(x= as.factor(0), y= mean.total_vel))+
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75))+
  geom_point()+
  theme_classic()
```

# Exploring a Loop for Visualizations 

```{r}

x <- as.factor(0)
 viz <- annas_clean %>%
   cross_join(x, copy = TRUE)
 
 colNames <- names(viz)[1:17]
 
 for (i in colNames) {
   plot <- ggplot(viz, aes_string(x= viz$y, y = i))+ 
     geom_violin(draw_quantiles = c(0.25, 0.5, 0.75))+
     geom_point() +
     theme_classic()
  print(plot)
  Sys.sleep(2)
   
 }

```

```{r}
normality_check <- as.data.frame(annas_clean)
columns <- names(normality_check)

for (i in columns) {
  plot <- hist(x = normality_check[,i])
  plot2 <- qqnorm(normality_check[,i]) 
    qqline(normality_check[,i])
  print(plot)
  print(plot2)
}

```

# Examining data distributions 
For each of the above distributions, I will group based on if the data is normal or other distributed. Then using the mean and variance from annas_summ data frame, I will be able to establish fake data for this data set upon which to build a model. 
**The list of variables, in order, are below. This order corresponds with the order of plots from the above loop.**
 1. mass
 2. massSTD
 3. w.lifted         
 4. w.liftedSTD      
 5. wing.ar          
 6. wing.arSTD       
 7. wing.length     
 8. wing.lengthSTD   
 9. mean.hor_acel   
 10. mean.hor_decel   
 11. mean.arc_avghvel 
 12. mean.total_vel   
 13.ss.pitch_roll   
 14. mean.pr_time    
 15. mean.pr_degrees 
 16. mean.arc_force   
 17. PRTpercent
 
 **Proposed Distributions**
**Normal**
3. w.lifted         
  4. w.liftedSTD 
5. wing.ar - appears to increase at end, hard to tell but most close to normal    
    6. wing.arSTD 
7. wing.length     
  8. wing.lengthSTD
16. mean.arc_force   

**Beta**
 1. mass - skewed right 
  2. massSTD
 
9. mean.hor_acel - left-skewed data
10. mean.hor_decel   - left skewed data
11. mean.arc_avghvel - left skewed data
12. mean.total_vel  - left skewed data
13.ss.pitch_roll - right skewed  
14. mean.pr_time - right skwewed 
15. mean.pr_degrees - left skewed data
17. PRTpercent - left skewed
 

```{r}

#seeing if body mass is normally distributed
hist(x= annas$mass, main= "Mass")
qqnorm(annas$mass, main='Mass')
qqnorm(annas$mass, main='Mass')
qqline(annas$mass)
```
appears that Mass is not normally distributed, based on right skew. Working idea: data could be represented by a Beta or Weibel distribution. 

```{r}
hist(x= annas$wing.length, main= "wing.length")
qqnorm(annas$wing.length, main='wing.length')
qqline(annas$wing.length)
```

```{r}
hist(x= annas$w.lifted, main= "w.lifted")
qqnorm(annas$w.lifted, main='w.lifted')
qqline(annas$w.lifted)
```
Weight lifted appears to be relatively normally distributed
Loop to explore data 
```{r}
for(i in 9:ncol(annas)){
  hist(x= annas[,i])
qqnorm(annas[,i])
qqline(annas[,i])
}
```


# Moment matching for mass data to see fit of Beta. 
```{r}
alpha <- mass_mu*(mass_mu*(1-mass_mu)/mass_var  -1)
beta <- (1-mass_mu)*(mass_mu*(1-mass_mu)/mass_var -1)

```

# Segre Notes
Below are notes from this paper and things to consider when modeling 
•	Birds that lifted more weight (accounting for wing morph) accelerated and decelerated faster, and did maneuvers with higher velocity. 
o	Executed faster, larger radius arcing turns
o	Centripetal acceleration of arcing turns not associated with burst capacity. 
o	Pitch-roll turns in less time, and for more of their heading changes (proportionally) 
•	Aspect ratio was a predictor for: centripetal acceleration and percent direction changes that were pitch-roll-turns. 
•	Body mass not predict in models (overlapped with zero). 
•	Competitor didn’t have substantial effect on performance metrics 
o	Horizontal acceleration and deceleration affected in opposite direction than expected.
o	Pitch down velocity increased 
o	More arcing turns used. 
•	When wings are predictors, only shape not size matters. 
•	Aspect ratio predicted some maneuverability, only for some maneuvers. 
•	Burst capacity is a brief increase in muscle strain and velocity: ability to do this should increase foraging behavior and competitive ability. 
•	Use of arcing and pitch roll turns the only metric influenced by all morphology, burst muscle capacity and competitor presence. 



